Influences on the access to higher education: Visualization and modeling for Chilean data

### 2021-08-18 Paper_Higher_Education.Rmd
Add files via upload

---
title: "Influences on the access to higher education: Visualization and modeling for Chilean data"
author: 'mail: olea.duran@uc.cl'
date: "10/08/2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
#Setup
options(scipen = 999)
rm(list=ls())
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(echo = F)
library(tidyverse)
library(data.table)
library(bit64)
library(readstata13)
library(haven)
library(dplyr)
library(readr)
library(grid)
library(gridExtra)
library(factoextra)
library(chilemapas)
library(kableExtra)
library(rsconnect)
library(knitr)
library(lemon)
library(ggspatial)
library(sf)
library(ggsflabel)
library(grid)
library(gridExtra)
library(ggalluvial)
options(digits = 3)
```

@@

bar = ggplot(data=x, aes(x=reorder(factor(Region),Total_Egresados),y=Total_Egresados, fill=Total_Egresados))  + coord_flip() + theme_minimal() + geom_bar(stat="identity") + labs(x="Region",y="Graduates") + theme(legend.position = "none", axis.title = element_text(size=17), axis.text.y = element_text(size=14), axis.text.x = element_text(size=14)) + geom_text(aes(label = Porc), position = position_dodge(width= 1), vjust= 0, hjust = -0.05, size = 6, color="black") + scale_y_continuous(expand = expansion(mult = c(0, .1)))+scale_fill_distiller(palette = "RdBu", direction=1,limits=c(0,920000),breaks = seq(0,920000,150000))

chile_1 = ggplot() +  geom_sf(data = x %>% filter(Zoom==1), aes(fill=Total_Egresados, geometry=geometry)) +  geom_sf_label_repel(data=x %>% filter(Zoom==1),aes(label = Region), nudge_x = -3, nudge_y = 0, seed = 10 ,colour="black", size=5) + coord_sf(xlim = c( -77.1,-65), ylim = c(-38, -17))
chile_1 = chile_1 + theme(legend.position="right", legend.box = "vertical") 
chile_1 = chile_1 + xlab("Longitude") + ylab("Latitude") + annotation_scale() + annotation_north_arrow(location='tr')
chile_1 = chile_1 + theme(plot.title = element_text(hjust = 0.5))  # + theme(legend.title=element_blank()) 
graf_total = chile_1 + theme_bw() + theme(legend.position = "none", axis.title = element_text(size=15)) +scale_fill_distiller(palette = "RdBu", direction=1, limits=c(0,920000), breaks = seq(0,920000,150000))

chile_2 = ggplot() +  geom_sf(data = x %>% filter(Zoom==0), aes(fill=Total_Egresados, geometry=geometry)) +  geom_sf_label_repel(data=x %>% filter(Zoom==0),aes(label = Region), nudge_x = 3, nudge_y = 0, seed = 10 ,colour="black", size=5) + coord_sf(xlim = c( -77.1,-65), ylim = c(-57, -38))
chile_2 = chile_2 + theme(legend.position="right", legend.box = "vertical") + labs(fill="Graduates")
chile_2 = chile_2 + xlab("Longitude") + ylab("Latitude") + annotation_scale() + annotation_north_arrow(location='tr')
chile_2 = chile_2 + theme(plot.title = element_text(hjust = 0.5),axis.title = element_text(size=15), legend.text = element_text(size=15)) +scale_fill_distiller(palette = "RdBu", direction=1,limits=c(0,920000), breaks = seq(0,920000,150000))  # + theme(legend.title=element_blank()) 
graf_total2 = chile_2 + theme_bw() + theme(legend.position = "right", axis.title = element_text(size=15), legend.text = element_text(size=12), legend.title = element_text(size=15)) +scale_fill_distiller(palette = "RdBu",direction=1,limits=c(0,920000), breaks = seq(0,920000,150000))

@@

todos <- colegios %>% group_by(AGNO_EGRESO) %>% summarise(No_Matriculado=sum(No_Matriculado),Matriculado=sum(Matriculado)) %>% ungroup() %>% mutate(Tasa=Matriculado/(Matriculado+No_Matriculado)*100)
todos$COD_DEPE2="All"

final <- bind_rows(municipales,subv,pagados,todos)
final <- final %>% mutate(AGNO_EGRESO=AGNO_EGRESO+1) #Admission Process is 1 + Graduation Year
final <- final %>% ungroup() %>% mutate(Porc = paste0(round(Tasa,1),"%"))

@@

Fig_5 = ggplot(final, aes(x=AGNO_EGRESO, y=Tasa, group=factor(COD_DEPE2))) +
  geom_line(aes(color=factor(COD_DEPE2))) +
  geom_point(aes(color=factor(COD_DEPE2)))+ 
  labs(x="Admission Process", y="Enrolment Rate", color="Dependence") +
  coord_cartesian(xlim = c(2008,2020.2), ylim = c(0,90))+ scale_x_continuous(breaks = c(2007:2020)) +
  scale_y_continuous(breaks = seq(0,90,10), sec.axis = dup_axis()) +
  theme_bw() +
  geom_rect(aes(xmin = 2007, xmax = 2010.5, ymin = -Inf, ymax = Inf),fill = "red", alpha = 0.005) +
  geom_rect(aes(xmin = 2010.5, xmax = 2014.5, ymin = -Inf, ymax = Inf),fill = "blue", alpha = 0.005) + 
  geom_rect(aes(xmin = 2014.5, xmax = 2018.5, ymin = -Inf, ymax = Inf),fill = "red", alpha = 0.005) +
  geom_rect(aes(xmin = 2018.5, xmax = 2020.8, ymin = -Inf, ymax = Inf),fill = "blue", alpha = 0.005) + 
  theme(plot.title = element_text(size = 12, face = "bold"),legend.title=element_text(size=16, angle=0),
        legend.text=element_text(size=14, angle = 0), axis.title.x = element_text(size = 18, vjust = -0.5), 
        axis.text.y = element_text(size=13, angle = 0), axis.title.y = element_text(size=18, angle=90), 
        axis.text.x = element_text(size=14, angle=90), axis.title.y.right = element_blank(), 
        legend.position = "top") + 
  geom_vline(xintercept = 2015.8, linetype="dashed") + 
  geom_text(aes(x=2015.6, label="Gratuidad", y=20), colour="blue", angle=90, text=element_text(size=1)) + 
  geom_vline(xintercept = 2011.8, linetype="dashed") +
  geom_text(aes(x=2011.6, label="Students' Movement", y=20), colour="red", angle=90, text=element_text(size=1)) + 
  geom_vline(xintercept = 2009.8, linetype="dashed") + 
  geom_text(aes(x=2009.6, label="LGE", y=20), colour="blue", angle=90, text=element_text(size=1)) + 
  geom_vline(xintercept = 2006.8) + 
  geom_vline(xintercept = 2019.8, linetype="dashed") + 
  geom_text(aes(x=2019.6, label="Social Outbreak", y=20), colour="red", angle=90, text=element_text(size=1)) +
  geom_text(aes(label = Porc), vjust= -0.5, hjust = 0, size = 4, color="black", check_overlap = TRUE, nudge_y = 0.5)

@@

final <- final %>% mutate(Total_Egresados = Matriculado+ No_Matriculado)
final <- final %>% filter(COD_DEPE2 !="All")
final <- final %>% ungroup() %>%  mutate(laggy = dplyr::lag(Total_Egresados, n = 1, default = NA))  
final <- final %>% mutate(Porc = ifelse(AGNO_EGRESO>2008, Total_Egresados/laggy*100-100,NA))
final <- final %>% ungroup() %>% mutate(Porc = ifelse(AGNO_EGRESO>2008,paste0(round(Porc,1),"%"),NA))

@@

Fig_2 = ggplot(final, aes(x=AGNO_EGRESO, y=Total_Egresados, group=factor(COD_DEPE2))) +
  geom_line(aes(color=factor(COD_DEPE2)))+
  geom_point(aes(color=factor(COD_DEPE2)))+ 
  labs(x="Admission Process", y="Graduates", color="Dependence") +
  coord_cartesian(xlim = c(2008,2020.2), ylim = c(0,100000)) +
  scale_x_continuous(breaks = c(2007:2020)) + 
  scale_y_continuous(breaks = seq(0,100000,20000), sec.axis = dup_axis()) + 
  theme_bw() +
  geom_rect(aes(xmin = 2007, xmax = 2010.5, ymin = -Inf, ymax = Inf),fill = "red", alpha = 0.005) + 
  geom_rect(aes(xmin = 2010.5, xmax = 2014.5, ymin = -Inf, ymax = Inf),fill = "blue", alpha = 0.005) + 
  geom_rect(aes(xmin = 2014.5, xmax = 2018.5, ymin = -Inf, ymax = Inf),fill = "red", alpha = 0.005) +
  geom_rect(aes(xmin = 2018.5, xmax = 2020.8, ymin = -Inf, ymax = Inf),fill = "blue", alpha = 0.005) + 
  theme(axis.title.y.right = element_blank(),plot.title = element_text(size = 12, face = "bold"),
        legend.title=element_text(size=16, angle=0),legend.text=element_text(size=14, angle = 0),
        axis.title.x = element_text(size = 18, vjust=-0.5),   axis.text.y = element_text(size=15, angle = 0),
        axis.title.y = element_text(size=18, angle=90), axis.text.x = element_text(size=14, angle=90),
        legend.position = "top") +
  geom_vline(xintercept = 2015.8) + 
  geom_text(aes(x=2015.6, label="Gratuidad", y=40000), colour="blue", angle=90, text=element_text(size=1)) + 
  geom_vline(xintercept = 2011.8) + 
  geom_text(aes(x=2011.6, label="St.Movement", y=40000), colour="red", angle=90, text=element_text(size=1)) +
  geom_vline(xintercept = 2009.8) + 
  geom_text(aes(x=2009.6, label="LGE", y=40000), colour="blue", angle=90, text=element_text(size=1)) + 
  geom_vline(xintercept = 2006.8) + 
  geom_vline(xintercept = 2019.8) + 
  geom_text(aes(x=2019.6, label="Soc.Outbreak", y=40000), colour="red", angle=90, text=element_text(size=1)) +
  geom_text(aes(label = Porc), vjust= -0.5, hjust = 0, size = 4, color="black", check_overlap = TRUE, nudge_y = 0.5)

@@

Fig_3 <- ggplot(final,aes(x = AGNO_EGRESO, y = Porcentaje, width = ancho, fill = COD_DEPE2)) +
  geom_col(colour = "black") +
  geom_text(aes(label = Porc), position = position_stack(vjust = 0.5)) + # if labels are desired
  facet_grid(~AGNO_EGRESO, scales = "free_x", space = "free_x") +
  scale_fill_brewer(palette = "Pastel1") +   
  theme_bw() + 
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        legend.position = "left", axis.title = element_blank()) +
   theme(legend.title=element_text(size=16, angle=0),legend.text=element_text(size=14, angle = 0),
        legend.position = "top", strip.text.x = element_text(size=14)) +
  labs(fill="Dependence")

@@

Fig_4 = ggplot(data=estudiantes, aes(x=COD_DEPE2, y=NEM, fill=factor(GEN_ALU))) +
  geom_split_violin() + 
  theme_bw() + 
  theme(plot.title = element_text(size = 12, face = "bold"),legend.title=element_text(size=16, angle=0),
        legend.text=element_text(size=14, angle = 0), axis.title.x = element_text(size = 18, vjust = -0.5), 
        axis.text.y = element_text(size=13, angle = 0), axis.title.y = element_text(size=18, angle=90), 
        axis.text.x = element_text(size=14, angle=0), axis.title.y.right = element_blank(), 
        legend.position = "top") + 
  labs(x="Dependence",y="HGPA", fill="Sex") + 
  geom_boxplot(width=0.2,color="white",alpha=0.2) + 
  theme(legend.position = "top")

@@

```{r Fig: 14, echo=FALSE, fig.align="center", fig.height=5, fig.width=10, message=FALSE, warning=FALSE}
setwd("~/Documents/GitHub/HE_Chile")
db_final <- read_rds("BASE_PAPER.RDS")
db_final <- db_final %>% filter(JOVEN=="1")
db_final <- db_final %>% mutate(COD_ENSE=ifelse(COD_ENSE<400,"H-C","T-P")) #%>% filter(COD_ENSE=="H-C")
names(db_final) <- toupper(names(db_final))
db_final <- db_final %>% mutate(AGNO_EGRESO = AGNO_EGRESO+1)
colegios <- db_final %>% group_by(AGNO_EGRESO, COD_DEPE2,PERCENTIL,MATRICULADO) %>% summarise(N=n())

colegios = dcast(colegios,AGNO_EGRESO + COD_DEPE2 + PERCENTIL ~ MATRICULADO)
names(colegios)[4] = "No_Matriculado";names(colegios)[5] = "Matriculado"

@@

colegios <- colegios %>% ungroup() %>% mutate(Porc = paste0(round(Tasa,1),"%"))

Fig_14 = ggplot(colegios, aes(x=AGNO_EGRESO, y=Tasa, group=factor(COD_DEPE2))) +
  geom_line(aes(color=factor(COD_DEPE2))) +
  geom_point(aes(color=factor(COD_DEPE2))) + 
  labs(x="Admission Process", y="Enrolment Rate", color="Dependence") +
  coord_cartesian(xlim = c(2008,2020.6), ylim = c(15,92))+
  scale_x_continuous(breaks = seq(2007,2020,1)) + 
  scale_y_continuous(breaks = seq(0,90,10), sec.axis = dup_axis()) + 
  theme_bw() +
  theme(plot.title = element_text(size = 12, face = "bold"),legend.title=element_text(size=16, angle=0),
        legend.text=element_text(size=14, angle = 0),axis.title.x = element_text(size = 18, vjust = -0.5), 
        axis.text.y = element_text(size=12, angle = 0),axis.title.y = element_text(size=18, angle=90), 
        axis.text.x = element_text(size=14, angle=90), axis.title.y.right = element_blank(), strip.text = element_text(size = 20),
        legend.position="top") +
  geom_text(aes(label = Porc), vjust= -0.5, hjust = 0, size = 4, color="black", check_overlap = TRUE, nudge_y = 0.5) + 
  facet_rep_wrap(~PERCENTIL, nrow=2,repeat.tick.labels = T)

Fig_14

ggsave("/Users/bolea/Documents/GitHub/HE_Chile/Figures/Fig_14.eps",Fig_14, width=23.67, height=12.6, limitsize = FALSE)
ggsave("/Users/bolea/Documents/GitHub/HE_Chile/Figures/Fig_14.pdf",Fig_14, width=23.67, height=12.6, limitsize = FALSE)

@@

Fig_6 = ggplot(colegios, aes(x=AGNO_EGRESO, y=Tasa, group=factor(GEN_ALU))) +
  geom_line(aes(color=factor(GEN_ALU))) +
  geom_point(aes(color=factor(GEN_ALU))) +
  theme() + 
  labs(x="Admission Process", y="Enrolment Rate", color="Sex") +
  coord_cartesian(xlim = c(2008,2020), ylim = c(40,60))+ scale_x_continuous(breaks = seq(2007,2020,1)) +
  scale_y_continuous(breaks = seq(40,60,5), sec.axis = dup_axis()) + 
  theme_bw() +
  theme(plot.title = element_text(size = 12, face = "bold"),legend.title=element_text(size=16, angle=0),
        legend.text=element_text(size=14, angle = 0),axis.title.x = element_text(size = 18, vjust = -0.5),
        axis.text.y = element_text(size=12, angle = 0),axis.title.y = element_text(size=18, angle=90), 
        axis.text.x = element_text(size=14, angle=90), axis.title.y.right = element_blank(),
        legend.position="top") + 
   geom_text(aes(label = Porc), vjust= -0.5, hjust = 0, size = 4, color="black", check_overlap = TRUE, nudge_y = 0.5)

@@

Alluvial <- Alluvial %>% mutate(TIPO_INST_3=ifelse(is.na(TIPO_INST_3),"No-Matriculado",TIPO_INST_3))
Alluvial <- Alluvial %>% mutate(OECD_AREA=ifelse(is.na(TIPO_INST_3),"No-Matriculado",OECD_AREA))
Alluvial <- Alluvial %>% filter(MATRICULADO==1)
Alluvial <- Alluvial %>% group_by(COD_ENSE,COD_DEPE2,TIPO_INST_3) %>% summarise(N=n())

@@

Fig_7 <- ggplot(as.data.frame(Alluvial),
       aes(y = N, axis1 = COD_DEPE2, axis2=TIPO_INST_3)) +
  geom_alluvium(aes(fill = COD_DEPE2), width = 1/12) +
  geom_stratum(width = 1/12, fill = "black", color = "grey") +
  geom_label(stat = "stratum", aes(label = after_stat(stratum))) +
  scale_x_discrete(limits = c("Dependence","Institution"), expand = c(.05, .05)) +
  scale_fill_brewer(type = "qual", palette = "Set1") + labs(y="Graduates enrolled", fill="Dependence") +
  theme_bw()+ theme(legend.position = "top",axis.title.x = element_text(size = 20, vjust = -0.5),
        axis.text.y = element_text(size=18, angle = 0),axis.title.y = element_text(size=18, angle=90), 
        axis.text.x = element_text(size=20, angle=0), legend.title=element_text(size=18, angle=0),
        legend.text=element_text(size=16, angle = 0))

@@

```{r Fig: 11, echo=FALSE, fig.height=7, fig.width=15, message=FALSE, warning=FALSE}
Data_2019 <- read_rds("BASE_PAPER.RDS")
Data_2019 <- Data_2019 %>% filter(JOVEN=="1")
Data_2019 <- Data_2019 %>% rename(COD_REG_ALU=codigo_region,Region=nombre_region)

df <- Data_2019 %>% filter(!is.na(COD_REG_ALU)) %>% filter(COD_REG_ALU>0)

  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        legend.position = "left", axis.title = element_blank()) +
   theme(legend.title=element_text(size=16, angle=0),legend.text=element_text(size=14, angle = 0),
        legend.position = "top", strip.text.x = element_text(size=14))

```{r Fig: 12 and 13, echo=FALSE, fig.height=7, fig.width=15, message=FALSE, warning=FALSE}

Data_2019 <- read_rds("BASE_PAPER.RDS")
Data_2019 <- Data_2019 %>% filter(JOVEN=="1")
Data_2019 <- Data_2019 %>% rename(COD_REG_ALU=codigo_region,Region=nombre_region)

#2007-2011 #remember admission process = agno_egreso + 1

>>>>

---
title: "Influences on the access to higher education: Visualization and modeling for Chilean data"
author: 'Bastian Olea-Duran, Francisco A. Segovia, Nixon Jerez-Lillo, Eduardo Alarcón-Bustamante, Danilo Alvares'
date: "10/08/2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
#Setup
options(scipen = 999)
rm(list=ls())
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(echo = F)
library(tidyverse)
library(data.table)
library(bit64)
library(readstata13)
library(haven)
library(dplyr)
library(readr)
library(grid)
library(gridExtra)
library(factoextra)
library(chilemapas)
library(kableExtra)
library(rsconnect)
library(knitr)
library(lemon)
library(ggspatial)
library(sf)
library(ggsflabel)
library(grid)
library(gridExtra)
library(ggalluvial)
library(ggrepel)
options(digits = 3)
```

@@

bar = ggplot(data=x, aes(x=reorder(factor(Region),Total_Egresados),y=Total_Egresados, fill=Total_Egresados))  + 
  coord_flip() + theme_minimal() + geom_bar(stat="identity") + labs(x="Region",y="Graduates") + 
  theme(legend.position = "none", axis.title = element_text(size=17), axis.text.y = element_text(size=14), 
        axis.text.x = element_text(size=14)) + 
  geom_text(aes(label = Porc), position = position_dodge(width= 1), vjust= 0, hjust = -0.05, size = 6, color="black") +
  scale_y_continuous(expand = expansion(mult = c(0, .3))) + 
  scale_fill_distiller(palette = "RdBu", direction=1, limits=c(13000,920000), breaks = seq(13000,920000,150000))

chile_1 = ggplot() +  geom_sf(data = x %>% filter(Zoom==1), aes(fill=Total_Egresados, geometry=geometry)) +  geom_sf_label_repel(data=x %>% filter(Zoom==1),aes(label = Region), nudge_x = -3, nudge_y = 0, seed = 10 ,colour="black", size=5) + coord_sf(xlim = c( -77.1,-65), ylim = c(-38, -17))
chile_1 = chile_1 + theme(legend.position="right", legend.box = "vertical") 
chile_1 = chile_1 + xlab("Longitude") + ylab("Latitude") + annotation_scale() + annotation_north_arrow(location='tr')
chile_1 = chile_1 + theme(plot.title = element_text(hjust = 0.5))  # + theme(legend.title=element_blank()) 
graf_total = chile_1 + theme_bw() + theme(legend.position = "none", axis.title = element_text(size=15)) +
  scale_fill_distiller(palette = "RdBu", direction=1, limits=c(13000,920000), breaks = c(13000,300000,600000,900000))

chile_2 = ggplot() +  geom_sf(data = x %>% filter(Zoom==0), aes(fill=Total_Egresados, geometry=geometry)) +  geom_sf_label_repel(data=x %>% filter(Zoom==0),aes(label = Region), nudge_x = 3, nudge_y = 0, seed = 10 ,colour="black", size=5) + coord_sf(xlim = c( -77.1,-65), ylim = c(-57, -38))
chile_2 = chile_2 + theme(legend.position="right", legend.box = "vertical") + labs(fill="Graduates")
chile_2 = chile_2 + xlab("Longitude") + ylab("Latitude") + annotation_scale() + annotation_north_arrow(location='tr')
chile_2 = chile_2 + theme(plot.title = element_text(hjust = 0.5),axis.title = element_text(size=15), legend.text = element_text(size=15)) +scale_fill_distiller(palette = "RdBu", direction=1,limits=c(0,920000), breaks = seq(0,920000,150000))  # + theme(legend.title=element_blank()) 
graf_total2 = chile_2 + theme_bw() + theme(legend.position = "right", axis.title = element_text(size=15), legend.text = element_text(size=12), legend.title = element_text(size=15)) +scale_fill_distiller(palette = "RdBu",direction=1,limits=c(13000,920000), breaks = c(13000,300000,600000,900000))

@@

final <- bind_rows(municipales,subv,pagados)
final <- final %>% mutate(AGNO_EGRESO=AGNO_EGRESO+1) #Admission Process is 1 + Graduation Year
final <- final %>% ungroup() %>% mutate(Porc = paste0(round(Tasa,1),"%")) %>%
  mutate(COD_DEPE2 = factor(COD_DEPE2, levels=c("Public", "Semi-Private", "Private")))

@@

Fig_5 = ggplot(final, aes(x=AGNO_EGRESO, y=Tasa, group=factor(COD_DEPE2))) +
  geom_line(linetype = "3313")+
  geom_point(aes(shape=factor(COD_DEPE2)), size=4)+ 
  labs(x="Admission Process", y="Enrolment Rate", shape="Dependence") +
  coord_cartesian(xlim = c(2008,2020.2), ylim = c(0,90))+ scale_x_continuous(breaks = c(2007:2020)) +
  scale_y_continuous(breaks = seq(0,90,10), sec.axis = dup_axis()) +
  theme_bw() +
  theme(plot.title = element_text(size = 12, face = "bold"),legend.title=element_text(size=16, angle=0),
        legend.text=element_text(size=14, angle = 0), axis.title.x = element_text(size = 18, vjust = -0.5), 
        axis.text.y = element_text(size=13, angle = 0), axis.title.y = element_text(size=18, angle=90), 
        axis.text.x = element_text(size=14, angle=90), axis.title.y.right = element_blank(), 
        legend.position = "top") + 
  geom_vline(xintercept = 2015.8, linetype="dashed") + 
  annotate(geom = "text", x = 2015.6, y = 20, label = "Gratuidad", color = "black", angle = 90) + 
  geom_vline(xintercept = 2011.8, linetype="dashed") +
  annotate(geom = "text", x = 2011.6, y = 20, label = "St.Movement", color = "black", angle = 90) + 
  geom_vline(xintercept = 2009.8, linetype="dashed") + 
  annotate(geom = "text", x = 2009.6, y = 20, label = "LGE", color = "black", angle = 90) + 
  geom_vline(xintercept = 2006.8) + 
  geom_vline(xintercept = 2019.8, linetype="dashed") + 
  annotate(geom = "text", x = 2019.6, y = 20, label = "Soc.Outbreak", color = "black", angle = 90) + 
  geom_text(aes(label = Porc), vjust= -0.9, hjust = 0, size = 4, color="black", check_overlap = TRUE, nudge_y = 0.5) +
    scale_shape_manual(values = c('Public'=15, 'Semi-Private'=17, 'Private'=16))

@@

final <- final %>% mutate(Total_Egresados = Matriculado+ No_Matriculado)
final <- final %>% filter(COD_DEPE2 !="All")
final <- final %>% ungroup() %>%  mutate(laggy = dplyr::lag(Total_Egresados, n = 1, default = NA))  
final <- final %>% mutate(Porc = ifelse(AGNO_EGRESO>2008, Total_Egresados/laggy*100-100,NA))
final <- final %>% ungroup() %>% mutate(Porc = ifelse(AGNO_EGRESO>2008,paste0(round(Porc,1),"%"),NA)) %>%
  mutate(COD_DEPE2 = factor(COD_DEPE2, levels=c("Public", "Semi-Private", "Private")))

@@

Fig_2 = ggplot(final, aes(x=AGNO_EGRESO, y=Total_Egresados, group=factor(COD_DEPE2))) +
  geom_line(linetype = "3313")+
  geom_point(aes(shape=factor(COD_DEPE2)), size=4)+ 
  labs(x="Admission Process", y="Graduates", shape="Dependence") +
  coord_cartesian(xlim = c(2008,2020.2), ylim = c(0,100000)) +
  scale_x_continuous(breaks = c(2007:2020)) + 
  scale_y_continuous(breaks = seq(0,100000,20000), sec.axis = dup_axis()) + 
  theme_bw() +
  theme(axis.title.y.right = element_blank(),plot.title = element_text(size = 12, face = "bold"),
        legend.title=element_text(size=16, angle=0),legend.text=element_text(size=14, angle = 0),
        axis.title.x = element_text(size = 18, vjust=-0.5),   axis.text.y = element_text(size=15, angle = 0),
        axis.title.y = element_text(size=18, angle=90), axis.text.x = element_text(size=14, angle=90),
        legend.position = "top") +
  geom_vline(xintercept = 2015.8, linetype="dashed") + 
  annotate(geom = "text", x = 2015.6, y = 40000, label = "Gratuidad", color = "black", angle = 90) + 
  geom_vline(xintercept = 2011.8, linetype="dashed") + 
  annotate(geom = "text", x = 2011.6, y = 40000, label = "St.Movement", color = "black", angle = 90) + 
  geom_vline(xintercept = 2009.8, linetype="dashed") + 
  annotate(geom = "text", x = 2009.6, y = 40000, label = "LGE", color = "black", angle = 90) + 
  geom_vline(xintercept = 2006.8, linetype="dashed") + 
  geom_vline(xintercept = 2019.8, linetype="dashed") + 
  annotate(geom = "text", x = 2019.6, y = 40000, label = "Soc.Outbreak", color = "black", angle = 90) + 
  geom_text(aes(label = Porc), vjust= -0.9, hjust = 0, size = 4, color="black", check_overlap = TRUE, nudge_y = 0.5) +
  scale_shape_manual(values = c('Public'=15, 'Semi-Private'=17, 'Private'=16))

@@

Fig_3 <- ggplot(final,aes(x = AGNO_EGRESO, y = Porcentaje, width = ancho, fill = COD_DEPE2)) +
  geom_col(colour = "black") +
  geom_text(aes(label = Porc), position = position_stack(vjust = 0.5)) + # if labels are desired
  facet_grid(~AGNO_EGRESO, scales = "free_x", space = "free_x") +
  scale_fill_brewer(palette = "Pastel1") +  
  theme_bw() + 
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        legend.position = "left", axis.title = element_blank()) +
   theme(legend.title=element_text(size=16, angle=0),legend.text=element_text(size=14, angle = 0),
        legend.position = "top", strip.text.x = element_text(size=14)) +
  labs(fill="Dependence") + scale_fill_grey(start=0.95, end=0.6)

@@

estudiantes <- estudiantes %>% mutate(COD_DEPE2 = factor(COD_DEPE2, levels=c("Public", "Semi-Private", "Private")))

Fig_4 = ggplot(data=estudiantes, aes(x=COD_DEPE2, y=NEM, fill=factor(GEN_ALU))) +
  geom_split_violin() + 
  theme_bw() + scale_fill_grey(start = 0.5, end = 0.7) +
  theme(plot.title = element_text(size = 12, face = "bold"),legend.title=element_text(size=16, angle=0),
        legend.text=element_text(size=14, angle = 0), axis.title.x = element_text(size = 18, vjust = -0.5), 
        axis.text.y = element_text(size=13, angle = 0), axis.title.y = element_text(size=18, angle=90), 
        axis.text.x = element_text(size=15, angle=0), axis.title.y.right = element_blank(), 
        legend.position = "top") + 
  labs(x="Dependence",y="HGPA", fill="Sex") + 
  geom_boxplot(width=0.2,color="white",alpha=0.2) + 
  theme(legend.position = "top")

@@

```{r Fig: 11, echo=FALSE, fig.align="center", fig.height=5, fig.width=10, message=FALSE, warning=FALSE}
setwd("~/Documents/GitHub/HE_Chile")
db_final <- read_rds("BASE_PAPER.RDS")
db_final <- db_final %>% filter(JOVEN=="1")
db_final <- db_final %>% mutate(COD_ENSE=ifelse(COD_ENSE<400,"H-C","T-P")) #%>% filter(COD_ENSE=="H-C")
names(db_final) <- toupper(names(db_final))
db_final <- db_final %>% mutate(AGNO_EGRESO = AGNO_EGRESO+1,
                                PERCENTIL = ifelse(PERCENTIL<=50, "ClassRank 1-50%","ClassRank 51-100%" ))

colegios <- db_final %>% group_by(AGNO_EGRESO, COD_DEPE2,PERCENTIL,MATRICULADO) %>% summarise(N=n())

table(colegios$PERCENTIL)
colegios = dcast(colegios,AGNO_EGRESO + COD_DEPE2 + PERCENTIL ~ MATRICULADO)
names(colegios)[4] = "No_Matriculado";names(colegios)[5] = "Matriculado"

@@

colegios <- colegios %>% ungroup() %>% mutate(Porc = paste0(round(Tasa,1),"%"))
colegios <- colegios %>%  mutate(COD_DEPE2 = factor(COD_DEPE2, levels=c("Public", "Semi-Private", "Private")))

Fig_11 = ggplot(colegios, aes(x=AGNO_EGRESO, y=Tasa, group=factor(COD_DEPE2))) +
  geom_line(linetype = "3313")+
  geom_point(aes(shape=factor(COD_DEPE2)), size=3)+ 
  labs(x="Admission Process", y="Enrolment Rate", shape="Dependence") +
  coord_cartesian(xlim = c(2008,2020.6), ylim = c(15,92))+
  scale_x_continuous(breaks = seq(2007,2020,1)) + 
  scale_y_continuous(breaks = seq(0,90,10), sec.axis = dup_axis()) + 
  theme_bw() +
  theme(plot.title = element_text(size = 12, face = "bold"),legend.title=element_text(size=16, angle=0),
        legend.text=element_text(size=14, angle = 0),axis.title.x = element_text(size = 18, vjust = -0.5), 
        axis.text.y = element_text(size=12, angle = 0),axis.title.y = element_text(size=18, angle=90), 
        axis.text.x = element_text(size=14, angle=90), axis.title.y.right = element_blank(), strip.text = element_text(size = 16),
        legend.position="top") +
  #geom_text(aes(label = Porc), vjust= -0.9, hjust = 0, size = 2, color="black", check_overlap = TRUE, nudge_y = 0.5) + 
  facet_rep_wrap(~PERCENTIL, nrow=1,repeat.tick.labels = T) +
  scale_shape_manual(values = c('Public'=15, 'Semi-Private'=17, 'Private'=16))

ggsave("/Users/bolea/Documents/GitHub/HE_Chile/Figures/Fig_11.eps",Fig_11, width=10.16, height=7.45, limitsize = FALSE)
ggsave("/Users/bolea/Documents/GitHub/HE_Chile/Figures/Fig_11.pdf",Fig_11, width=10.16, height=7.45, limitsize = FALSE)

@@

Fig_6 = ggplot(colegios, aes(x=AGNO_EGRESO, y=Tasa, group=factor(GEN_ALU))) +
  geom_line(linetype = "3313") +
  geom_point(aes(shape=factor(GEN_ALU)), size=4) +
  theme() + 
  labs(x="Admission Process", y="Enrolment Rate", shape="Sex") +
  coord_cartesian(xlim = c(2008,2020), ylim = c(40,60))+ scale_x_continuous(breaks = seq(2007,2020,1)) +
  scale_y_continuous(breaks = seq(40,60,5), sec.axis = dup_axis()) + 
  theme_bw() + scale_color_grey(start = 0, end = .7) +
  theme(plot.title = element_text(size = 12, face = "bold"),legend.title=element_text(size=16, angle=0),
        legend.text=element_text(size=14, angle = 0),axis.title.x = element_text(size = 18, vjust = -0.5),
        axis.text.y = element_text(size=12, angle = 0),axis.title.y = element_text(size=18, angle=90), 
        axis.text.x = element_text(size=14, angle=90), axis.title.y.right = element_blank(),
        legend.position="top") + 
   geom_text_repel(aes(label = Porc))+
   scale_shape_manual(values = c('Female'=16, 'Male'=17))

@@

Alluvial <- Alluvial %>% mutate(TIPO_INST_3=ifelse(is.na(TIPO_INST_3),"Not-Enrolled",TIPO_INST_3))
Alluvial <- Alluvial %>% mutate(OECD_AREA=ifelse(is.na(TIPO_INST_3),"Not-Enrolled",OECD_AREA))
#Alluvial <- Alluvial %>% filter(MATRICULADO==1)
Alluvial <- Alluvial %>% group_by(COD_ENSE,COD_DEPE2,TIPO_INST_3) %>% summarise(N=n())

@@

Alluvial <- Alluvial %>%  mutate(COD_DEPE2 = factor(COD_DEPE2, levels=c("Public", "Semi-Private", "Private")),
                                 TIPO_INST_3 = factor(TIPO_INST_3, levels = c("Not-Enrolled","CRUCH","Non-CRUCH","PI or TTC")))

Fig_7 <- ggplot(as.data.frame(Alluvial),
       aes(y = N, axis1 = COD_DEPE2, axis2=TIPO_INST_3)) +
  geom_alluvium(aes(fill = COD_DEPE2), width = 1/12) +
  geom_stratum(width = 1/12, fill = "black", color = "grey") +
  geom_label(stat = "stratum", aes(label = after_stat(stratum))) +
  scale_x_discrete(limits = c("Dependence","Institution"), expand = c(.05, .05)) +
  scale_fill_brewer(type = "qual", palette = "Set1") + labs(y="Graduates Enrolled", fill="Dependence") +
  theme_bw()+ theme(legend.position = "top",axis.title.x = element_text(size = 20, vjust = -0.5),
        axis.text.y = element_text(size=18, angle = 0),axis.title.y = element_text(size=18, angle=90), 
        axis.text.x = element_text(size=20, angle=0), legend.title=element_text(size=18, angle=0),
        legend.text=element_text(size=16, angle = 0))

@@

```{r Case of Schools - Fig: 8, 9, 10, 12, 13 and 14, echo=FALSE, fig.align="center", fig.height=8, fig.width=12, message=FALSE, warning=FALSE}
library(rpart)
library(dplyr)
library(readr)
library(rattle)
library(ggplot2)
library(reshape2)
library(ggforce)
library(gridExtra)
library(ggthemes)
setwd("~/Documents/GitHub/HE_Chile")
